---
title: "Practice Activity 5.2"
execute:
  echo: true
  error: true 
author: "Fiona Norton"
format: 
  html: 
    code-fold: true
    code-summary: "Show code"
    theme: "materia"
editor: visual
---

## Setup

```{r}
#| message: false
library(tidyverse)
library(tidyr)
library(stringr)

message <- read_csv("https://www.dropbox.com/s/lgpn3vmksk3ssdo/scrambled_message.txt?dl=1") %>% 
  pull(Word)
```

## Warm-up Exercises 

1. How many characters are in the scrambled message?
```{r}
str_length(message) |> 
  sum()
```

2. How many words are in the scrambled message?
```{r}
length(message)
```

3. Print out every piece of a word in the scrambled message that starts with the letter “m”.
```{r}
message |> 
  str_subset("^m")
```

4. Print out every piece of a word in the scrambled message that ends with the letter “z”
```{r}
message |> 
  str_subset("z$")
```

5. Print out every punctuation symbol in the scrambled message.
```{r}
message |> 
  str_subset("[:punct:]")
```

6. Print out the longest word in the scrambled message.
```{r}
message[str_length(message) == max(str_length(message))]
```

## Decode the Message

Complete the following steps to decode the message.

1. Remove any spaces before or after each word.
2. No word should be longer than 16 characters. Drop all extra characters off the end of each word.
3. Any time you see the word “ugh”, with any number of h’s, followed by a punctuation mark, delete this.
4. Replace all instances of exactly 2 a’s with exactly 2 e’s.
5. Replace all z’s with t’s.
6. Every word that ends in b, change that to a y.
7. Every word that starts with k, change that to a v.
8. Use str_c() to recombine all your words into a message.
9. Find the movie this quote is from.

```{r}
message <- message |> 
  str_trim("both") |> 
  str_trunc(16, "right") |> 
  str_replace_all("ug(h+)(?=[:punct:])", "") |> 
  str_replace_all("a{2}", "ee") |> 
  str_replace_all("z", "t") |> 
  str_replace_all("b$", "y") |> 
  str_replace_all("^k", "v") |> 
  str_flatten(collapse = " ")
```

```{r}
str_subset(message, "ug(h+)(?=[:punct:])")
```




