---
title: "Practice Activity 9 - Simulation"
execute:
  echo: true
  error: true 
author: "Fiona Norton"
format: 
  html: 
    code-fold: true
    code-summary: "Show code"
    theme: "journal"
editor: visual
---

# Background Info

```{r}
#| message: false
library(dplyr)
library(rlang)
library(purrr)
```


Trombones weights are Normally distributed, with a mean weight of 4.6 pounds and a standard deviation of 0.8 pounds.

Trumpet weights are Uniformly distributed between 1.5 and 3.5 pounds.

Reed instruments (like clarinets, bassoons, and saxophones) have weights that are Chi-Square distributed with 4 degrees of freedom.


# Warm-Up

What is the 95th percentile for trumpet weight?

```{r}
qunif(0.95, min = 1.5, max = 3.5)
```


What is the 10th percentile for trombone weight?

```{r}
qnorm(0.10, mean = 4.6, sd = 0.8)
```


About what percent of trombones do you expect to be more than 5 pounds?

```{r}
1 - pnorm(5, mean = 4.6, sd = 0.8)
```


About what percent of reed instruments do you expect to be more than 5 pounds?

```{r}
1 - pchisq(5, df = 4)
```


Simulate 100 random trombone weights. How many of them were below 4 pounds?

```{r}
sim_trombones <- rnorm(100, mean = 4.6, sd = 0.8)
sim_trombones < 4
```


# Catch a Con 

Items included:
    Trombones: 76
    Cornets: 110
    Saxophones: 542
    Clarinets: 318
    Bassoons: 175
    
Total Shipped Weight: 4532 lbs.

```{r}
music_man <- function(n_tromb, n_cor, n_reed){
  
  trombones <- rnorm(n_tromb, mean = 4.6, sd = 0.8)
  cornets <- runif(n_cor, min = 1.5, max = 3.5)
  reeds <- rchisq(n_reed, df = 4)
  
  return(sum(trombones, cornets, reeds))
  
}

set.seed(1957)

my_weights <- rerun(.n = 1000, 
                    music_man(n_tromb = 76, n_cor = 110, n_reed = (542 + 318 + 175))
                    ) |> 
  flatten_dbl()


```


```{r}
less_than_shipment <- my_weights < 4532
sum(less_than_shipment) / 1000
```


