---
title: "Lab 5: Factors in Visualization"
execute:
  echo: true
  error: true 
author: "Fiona Norton"
format: 
  html: 
    code-fold: true
    code-summary: "Show code"
    theme: "spacelab"
    self-contained: true
editor: visual
---

# Revisiting Lab 2

## Setup

```{r}
#| message: false
library(tidyverse)
library(here)
library(ggplot2)
surveys <- read_csv(here :: here("supporting_artifacts", "datasets", "surveys.csv"))
```

## Reorder the boxplots and add a title.

```{r}
ggplot(data = surveys, 
       mapping = aes(y = fct_reorder(species, weight), x = weight) 
       ) + 
  geom_boxplot() +
  geom_jitter(alpha = .2, color = "red") + 
  labs(x = "", y = "", title = "Species of Rodent by Weight (gm)") 
```

# Time Series Plot

## Create a visualization of how weights vary for each genus over the duration of the study.

```{r}
ggplot(data = surveys,
       mapping = aes(x = year, y = weight, color = genus)
       ) +
  geom_point() +
  facet_wrap(~genus) + 
  theme(legend.position = "none") +
  labs(x = "", y = "", title = "Weight (gm) of Each Genus of Rodent throughout the Study")
```

## Calculate and plot the mean weight for each year (for each genus).

```{r}
surveys |> 
  group_by(genus, year) |> 
  summarize(mean_weight = mean(weight)) |> 
  ggplot(mapping = aes(x = year, y = mean_weight, color = genus)) +
  geom_line() +
  labs(x = "", y = "", title = "Average Weight (gm) of Each Genus of Rodent throughout the Study")
```

## Reorder the lines so the weights go in descending order.

```{r}
surveys |> 
  group_by(genus, year) |> 
  summarize(mean_weight = mean(weight)) |> 
  ggplot(mapping = aes(x = year, y = mean_weight, color = fct_reorder2(genus, year, mean_weight))) +
  geom_line()
```

## Give the legend a new title and move the information from the y-axis to the plot title.

```{r}
surveys |> 
  group_by(genus, year) |> 
  summarize(mean_weight = mean(weight)) |> 
  ggplot(mapping = aes(x = year, y = mean_weight, color = fct_reorder2(genus, year, mean_weight))) +
  geom_line() +
  labs(x = "", y = "", 
       title = "Average Weight (gm) of Each Genus of Rodent throughout the Study (years)", 
       color = "Genus") 
```

# Captures over the week

## Create a visualization of the number of rodents captured each day of the week.

```{r}
surveys |> 
  ggplot(mapping = aes(x = day_of_week, fill = day_of_week)) +
  geom_bar() +
  labs(x = "", y = "", 
       title = "Number of Rodents Captured on Each Day of the Week") +
  theme(legend.position = "none") 
```

## Change the order of the day of the week to go Monday through Sunday.

```{r}
surveys |> 
  ggplot(mapping = aes(
    x = fct_relevel(day_of_week, c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")), 
    fill = day_of_week)) +
  geom_bar() +
  labs(x = "", y = "", 
       title = "Number of Rodents Captured on Each Day of the Week") +
  theme(legend.position = "none") 
```

## Plot the number of rodents captured between the week days and weekends.

```{r}
surveys |> 
  mutate(time_of_week = fct_collapse(day_of_week, 
                                     weekday = c("Mon", "Tue", "Wed", "Thu", "Fri"), 
                                     weekend = c("Sat", "Sun"))) |> 
  ggplot(mapping = aes(
    x = time_of_week, 
    fill = time_of_week)) +
  geom_bar() +
  labs(x = "", y = "", 
       title = "Number of Rodents Captured on Weekdays vs. Weekends") +
  theme(legend.position = "none")
           
```
